### 2022-11-07 paper-thesis/chapters/03-variable-capture.tex
Create paper-thesis
Create more structure

The implementation of the base language shown in section \ref{chap:baselang} has one big problem, that is name collisions. This section will explore several ways of solving these collisions. An example of such a collision is the following: What is the type of this expression (a polymorphic identity function)?
\begin{lstlisting}
\T : Type. \T : T. T
\end{lstlisting}

@@

Almost all compilers that typecheck dependently typed languages use de Bruijn representation for variables \cite{TODO agda lean etc}. Using de Bruijn indices in statix is possible, but sacrifices a lot. It would require a transformation on the AST before typechecking. Modifying the AST like this causes problems, because it changes AST nodes. All editor services that rely on \verb|.ref| annotations, such as renaming, can no longer be used. It also loses a lot of the benefits of using Spoofax, since using scope graphs relies on using names.

@@

\subsection{Improving the readability of types}

>>>>

The implementation of the base language shown in chapter \ref{chap:baselang} has one big problem, that is name collisions. This section will explore several ways of solving these collisions. An example of such a collision is the following: What is the type of this expression (a polymorphic identity function)?
\begin{lstlisting}
\T : Type. \T : T. T
\end{lstlisting}

@@

Almost all compilers that typecheck dependently typed languages use de Bruijn representation for variables. Using de Bruijn indices in statix is possible, but sacrifices a lot. It would require a transformation on the AST before typechecking. Modifying the AST like this causes problems, because it changes AST nodes. All editor services that rely on \verb|.ref| annotations, such as renaming, can no longer be used. It also loses a lot of the benefits of using Spoofax, since using scope graphs relies on using names.

@@

\section{Improving the readability of types}

Because the expression above, with \verb|ScopedName|s, is not particularly readable, we add a new post-analysis Stratego pass that converts the \verb|ScopedName|s to ticked named. For example, the above would be transformed to:

\begin{lstlisting}
FnType(T, Type(), FnType(T', Var(T)), Var(T)))
\end{lstlisting}
Ticks are added to names where necessary. We do this by following these rules:

\begin{enumerate}
	\item When we encounter a \verb|ScopedName| for the first time, we keep adding ticks to the name until we find a name that has not been used before. 
	\item We define a dynamic rule \verb|Rename :: string -> string| and we store the new name we generated using this rule.
	\item When we encounter a \verb|ScopedName|, use the \verb|Rename| rule to find what the name was transformed to.
\end{enumerate}
